Impact of the High Population Events on Urban Mobility: Who is Affected the Most?

### 2019-12-10 manuscript.Rmd
added bibliography

This paper explores mobility in Washington, D.C. during major events such as the 2017 women's march and inauguration. On January 20, 2017, Donald Trump was inaugurated in Washington. The following day, January 21, hundreds of thousands of people protested Trump's inauguration through the Women's March. The women's march in Washington has been called the largest protest in the United States with an estimated 470,000 people in attendance [@wallace_parlapiano_2017]. Crowd scientists estimate that the crowd at the Trump inauguration was about a third of the size, making it around 160,000 people [@wallace_parlapiano_2017]. Population increases of this size over the course of just one weekend creates a huge shock to the transportation system. Before modeling this impact, however, we first modeled the impact of demographic census data on travel time during normal January conditions. We then were able to observe how demographics impact mobility and how those factors are exacerbated during major shocks to the system. We found that the number of roads, distance to the Washington Monument, and the number of black residents are consistently the most significant predictors of travel time to each polygon from the Washington Monument. Further, we found that ________ ADD FINDINGS ABOUT SIGNIFICANCE DURING SHOCK!!!

@@

Many scholars have studied both mobility and shocks to the system in a variety of contexts. It is important to closely study every city's transportation system which is why many studies have been conducted. Noulas et. al use a network called Foursquare which records peoples locations to track and model mobility in a number of different cities [@noulas2012tale]. Studies about mobility and accessibility have been conducted in many cities including New York City [@litman2017evaluating] where concepts such as accessibility can be quantified and modeled to demonstrate lapses in the system. Shocks to transportation systems have also been modeled by a number of scholars. Donovan and Work's 2017 study uses taxi data to model the impacts of major events such as hurricane Sandy on New York city [@donovan2017empirically]. Wang also modeled the impact of Hurricane Sandy on New York city by looking at images of people throughout the city [@wang2014quantifying]. Other major events modeled by scholars include the Olympics [@friedman2001impact] and the World Cup [@de2014transportation]. Scholars have used a variety of data sources to observe the mobility of people including the taxi data and images as mentioned above, and cell phone data [@calabrese2010geography]. Uber data is accessible and also provides an indication of mobility since so many people travel in cars. Because of this accessibility, we decided to model the inauguration and Women's March in Washington, D.C., two major events that occurred back to back.

@@

travel_time_sf %>% filter(year == 2017, month == 1, week == 3) %>%
	ggplot() +
	geom_sf(mapping = aes(fill = black_pct)) +
	theme_map +
	scale_fill_viridis_c()

travel_time_sf %>% filter(year == 2017, month == 1, week == 3) %>%
	ggplot() +
	geom_sf(mapping = aes(fill = car_transit_pct)) +
	theme_map +
	scale_fill_viridis_c()
```

@@

```{r}
travel_time_sf_nontreat %>%
	ggplot() +
	geom_sf(mapping = aes(fill = mtt_mean)) +
	theme_map +
	scale_fill_viridis_c()

travel_time_sf_nontreat %>%
	ggplot() +
	geom_sf(mapping = aes(fill = mtt_sd)) +
	theme_map +
	scale_fill_viridis_c()

travel_time_sf_nontreat %>%
	ggplot() +
	geom_sf(mapping = aes(fill = log_roads)) +
	theme_map +
	scale_fill_viridis_c()

travel_time_sf_nontreat %>%
	ggplot() +
	geom_sf(mapping = aes(fill = log_transit)) +
	theme_map +
	scale_fill_viridis_c()
```

>>>>

Examining Washington D.C. during the 2017 women's march and presidential inauguration, this paper looks to examine the effect that large scale public events have on urban mobility. On January 20, 2017, Donald Trump was inaugurated in Washington. The following day, January 21, hundreds of thousands of people protested Trump's inauguration through the Women's March. The women's march in Washington has been called the largest protest in the United States with an estimated 470,000 people in attendance. Crowd scientists estimate that the crowd at the Trump inauguration was about around 160,000 people, a third of the size of the protest [@wallace_parlapiano_2017]. A population increase of this size over the course of just one weekend creates a huge shock to the transportation system. This can occur through the influx of people coming into the city as well as road blocks impeding traffic. This study looks at changes in the mean travel time to the center of D.C., with the center point defined as the Washington Monument. We chose this point because it was near the presidential inauguration and women's march. Before modeling this impact, however, we first modeled the impact of demographic census data on travel time during normal January conditions. We then were able to observe how demographics impact mobility and how those factors are exacerbated during major shocks to the system. We found that the number of roads, distance to the Washington Monument, and the number of black residents are consistently the most significant predictors of travel time to each polygon from the Washington Monument. Further, we found that ________ ADD FINDINGS ABOUT SIGNIFICANCE DURING SHOCK!!!

@@

Many scholars have studied both mobility and shocks to the system in a variety of contexts. It is important to closely study every city's transportation system which is why many studies have been conducted. Noulas et. al use a network called Foursquare which records peoples locations to track and model mobility in a number of different cities [@noulas2012tale]. Studies about mobility and accessibility have been conducted in many cities including New York City [@litman2017evaluating] where concepts such as accessibility can be quantified and modeled to demonstrate lapses in the system. Shocks to transportation systems have also been modeled by a number of scholars. Donovan and Work's 2017 study uses taxi data to model the impacts of major events such as hurricane Sandy on New York city [@donovan2017empirically]. Wang also modeled the impact of Hurricane Sandy on New York city by looking at images of people throughout the city [@wang2014quantifying]. Other major events modeled by scholars include the Olympics [@friedman2001impact] and the World Cup [@de2014transportation]. Transportation scholars have used a variety of data sources to observe the mobility of people including taxi data, satellite imagery, and cell phone data [@calabrese2010geography]. Uber data is accessible and also provides an indication of mobility since so many people travel in cars. Because of this accessibility, we decided to model the effects of the inauguration and Women's March in Washington, D.C. that happened during the third weekend of January 2017.

@@

travel_time_sf %>% filter(year == 2017, month == 1, week == 3) %>%
	ggplot() +
	geom_sf(mapping = aes(fill = black_pct)) +
	theme_map +
	scale_fill_viridis_c()
```

@@

```{r}
# travel_time_sf_nontreat %>%
# 	ggplot() +
# 	geom_sf(mapping = aes(fill = mtt_mean)) +
# 	theme_map +
# 	scale_fill_viridis_c()
# 
# travel_time_sf_nontreat %>%
# 	ggplot() +
# 	geom_sf(mapping = aes(fill = mtt_sd)) +
# 	theme_map +
# 	scale_fill_viridis_c()
# 
# travel_time_sf_nontreat %>%
# 	ggplot() +
# 	geom_sf(mapping = aes(fill = log_roads)) +
# 	theme_map +
# 	scale_fill_viridis_c()
# 
# travel_time_sf_nontreat %>%
# 	ggplot() +
# 	geom_sf(mapping = aes(fill = log_transit)) +
# 	theme_map +
# 	scale_fill_viridis_c()
```
